### 2021-02-10 source/tex/ladybug_adhesion.tex

%% LyX 2.3.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{achemso}
\usepackage[T2A,LGR,T1]{fontenc}
\usepackage{babel}
\usepackage{array}
\usepackage{textcomp}
\usepackage{url}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subscript}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
 {hyperref}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.

\title{Wetting of adhesive fluid controls insect adhesion in air and underwater}

\author{Pranav Sudersan}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\author{Michael Kappl}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\author{Bat-El Pinchasik}

\affiliation{School of Mechanical Engineering, Tel Aviv University, Tel Aviv-Yafo,
Israel}

\author{Hans-J\"{u}rgen Butt}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\author{Thomas Endlein}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\email{endlein01@mpip-mainz.mpg.de}

\phone{+49 6131 379 111}

\fax{+49 6131 379 310}

\DeclareRobustCommand{\greektext}{%
  \fontencoding{LGR}\selectfont\def\encodingdefault{LGR}}
\DeclareRobustCommand{\textgreek}[1]{\leavevmode{\greektext #1}}
\ProvideTextCommand{\~}{LGR}[1]{\char126#1}

\DeclareRobustCommand{\cyrtext}{%
  \fontencoding{T2A}\selectfont\def\encodingdefault{T2A}}
\DeclareRobustCommand{\textcyr}[1]{\leavevmode{\cyrtext #1}}

\newcommand{\lyxmathsym}[1]{\ifmmode\begingroup\def\b@ld{bold}
  \text{\ifx\math@version\b@ld\bfseries\fi#1}\endgroup\else#1\fi}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\SectionNumbersOn
%\setkeys{acs}{doi = true}
\setkeys{Gin}{width=\linewidth}
\usepackage{subcaption}

\makeatother

\begin{document}
\begin{abstract}
Insects like beetles can stick to various surfaces using hairy pads
on their legs. It was previously shown that these pads can even attach
underwater, presumably due to an air bubble trapped around the pad.
However, the bubble's contribution to adhesion via capillary force
remained unclear. To investigate the role of the bubble, we measured\emph{
in-vivo} underwater adhesion of a ladybug's pad, in the presence and
absence of the trapped bubble and compare it with its adhesion in
air. Our experiments reveal that on a hydrophobic substrate, even
without a bubble, the pad can show adhesion underwater comparable
to that in air. On a hydrophilic substrate, underwater adhesion is
significantly reduced, with or without a trapped bubble. To explain
these results, we develop a simple theoretical model for the net adhesion
of a hairy pad due to capillary forces. Our results demonstrate that
the wetting properties of the adhesive fluid determines the insect's
adhesion in both air and underwater conditions.
\end{abstract}

\section{Introduction}

The question of how insects and other small animals can walk on smooth
surfaces against gravity has fascinated scientists for at least the
past three centuries \cite{RN198,RN59}. We now know that such animals
are able to adhere by using specialized organs on their feet called
adhesive pads. These adhesive pads exist in a variety of types depending
on the animal, but are generally categorized into: 1) ``smooth pads''
found in ants\cite{RN201}, stick insects \cite{RN182}, etc. and
2) ``hairy pads'' seen in flies \cite{RN155}, geckos \cite{RN202}
and others. The hairy pads show: 1) compliance to rough surfaces due
to their lower effective modulus, 2) angle dependent adhesion due
to asymmetric hair geometry and 3) self-cleaning capability \cite{RN20},
which makes them suitable to adhere to most surfaces reversibly. Many
of these insects pads also secrete an adhesive fluid, as seen in flies
and ants \cite{RN201} (``wet adhesion''), while others such as
spiders and geckos rely on their dry hairy pads for attachment (``dry
adhesion''). In the ``wet adhesion'' case, fluid secretion can
enforce adhesion through surface tension and viscous forces \cite{RN71},
while, ``dry adhesion'' relies mostly on van der Waals forces \cite{RN202}.

Terrestrial beetles such as the dock beetle or the ladybug have hairy
pads consisting of a dense array of hair-like structures called setae.
The setae tips can be discoidal, spatula or pointed shaped, which
are distributed throughout the pad depending on sex or species\cite{RN19}.
Single seta force measurements revealed that discoidal shaped seta
shows larger pull-off force than spatula and pointed setae\cite{RN79},
illustrating the role of hair geometry in adhesion. The tip of each
seta secretes approximately one femtoliter of oily adhesive fluid
\cite{RN108}. The fluid's chemical composition is identified as a
mixture of mostly long chain hydrocarbons\cite{RN96} with traces
of triglycerides, fatty acids and cholesterol\cite{RN221,RN222}.
A recent study by Gernay et. al.\cite{RN77}, based on an elastocapillary
model, has been able to reasonably predict single seta adhesion forces
theoretically, confirming the dominant role of surface tension in
the ``wet adhesion'' of beetles.

While most of the studies on insect adhesion are done under natural
conditions in air, insect attachment underwater has been relatively
unexplored. Typically, underwater adhesion is complicated to achieve
due to the difficulty in displacing the water layer and enable good
contact\cite{RN123}. Regardless, a study on leaf beetles\cite{RN87}
has revealed that they can in fact attach quite well to surfaces underwater.
Its hairy pad traps an air bubble underwater, which dewets the surface
on contact. It has been hypothesized that a combination of capillary
forces due the air bubble and hair secretions within the dewetted
area results in its adhesion underwater. However, a detailed investigation
of the bubble's contribution and necessity to adhere to different
surfaces is lacking. Geckos are also known to adhere underwater, where
its shear adhesion force on hydrophobic substrates are similar in
both air and underwater conditions. Interestingly, its adhesion on
a fluorinated substrate is even larger underwater than in air\cite{RN15}.
This has been partially explained by a thermodynamic work of adhesion
model, assuming full displacement of water at the interface, leading
to a dry ``van der Waals'' contact of hairs with the surface\cite{RN199}.

The goal of this paper is to provide a generalized picture of adhesion
for insects which use hairy pads and secrete a oily fluid for attachment.
First, we measured adhesion of a single constrained pad of a live
ladybug beetle in air and underwater conditions, both on smooth hydrophilic
and hydrophobic glass surfaces, with a microscopic observation of
the contact process. Second, we developed a simple theoretical model
considering capillary forces to predict the net adhesion force of
a hairy pad under different conditions. The case of underwater adhesion
was studied both in the presence and absence of a trapped bubble,
to decouple the bubble's role in the insect's adhesion. Finally, we
discuss key insights gained from our experiments and model with regards
to understanding adhesion in other animals.

\section{Experimental}

Normal adhesion force measurements on a restrained leg of a live ladybug
beetle were performed. We characterize adhesion by the pull-off force
during detachment. Measurements were done against smooth glass and
fluorinated surfaces to represent hydrophilic and hydrophobic substrates
respectively. When no water was present, we labeled the contact type
as ``\emph{in air}''. For the underwater conditions, measurements
were done both in the presence and absence of a trapped bubble (``\emph{underwater:
bubble}'' and ``\emph{underwater: no bubble}'', respectively) to
investigate the bubble's role in underwater adhesion. Adhesion force
for each of the labeled contact types were compared for both substrates.

\subsection{Material and Methods}

\subsubsection{Insect preparation}

Seven-spotted male adult ladybug beetles (\emph{Coccinella septempuctata})
purchased from Katz Biotech (Baruth, Germany) were used for adhesion
tests. The beetles were housed in a plastic box filled with leaves,
twigs and stones at room temperature and 60-80\% relative humidity
with daily access to sunlight. They were fed with raisins, honey and
water\emph{ ad libitum}.

Each leg of the beetle has a pair of hairy pads covered with mostly
discoidal shaped setae, capable of strong adhesion. For the test,
we only allowed the distal-end pad to make contact. The beetle's leg
was constrained similar to the method described by Bullock et al \cite{RN19}.
A steel ball fixed with a piece of thick solder wire served as a rotatable
holder to fix the beetle and align its leg (Figure \ref{fig:Setup}).
The beetle was first anesthetized using CO\textsubscript{2} sublimating
from a piece of dry ice and then glued to the steel ball on its back.
Its front left leg was carefully fixed to the solder wire of the holder
using Blue-Tac. Its claws were fixed using epoxy glue to prevent any
wiggling. A small piece of non-sticky Teflon tape helped to keep its
other legs tucked close to the body and avoid their interference during
the adhesion test.

After measurements, the beetle was freed by carefully removing the
epoxy glue and Blue-Tac without harming it and set free.

\subsubsection{Substrate preparation}

Standard 20 mm wide glass coverslips were used as the hydrophilic
substrate. Glass was wiped with isopropanol, rinsed in water and dried
under nitrogen flow. The surface was then plasma cleaned in a oxygen
plasma chamber (Diener Electronic Femto) for 10 min at 120 W. The
surface was further rinsed with water and dried under nitrogen flow.

For the hydrophobic substrate, the glass cover slip was coated with
a fluorosilane via chemical vapor deposition (CVD). First, the glass
was cleaned using IPA and plasma treated as before. Next, 0.2 ml of
Trichloro(1H,1H,2H,2H-perfluorooctyl) silane (PFOTS), procured from
Sigma Aldrich, was put in a sealed chamber along with the the cleaned
glass. The chamber was placed under 100 mbar pressure for 10 min for
the CVD process. Finally, the substrate was annealed at 150\textcelsius{}
for 3 hours.

The substrate wettability was characterized by dynamic contact angle
measurements, performed with a DataPhysics OCA 35 contact angle goniometer.
Milli-Q water and n-hexadecane were used as test liquids. Advancing
and receding contact angles were measured for a maximum drop volume
of 10 \textgreek{m}l and with 0.5 \textgreek{m}l/s flow rate.

\begin{table}[H]
\centering{}%
\begin{tabular}{|c|c|c|c|}
\hline 
Substrate & Liquid & \ensuremath{\theta}\textsubscript{A} & \ensuremath{\theta}\textsubscript{R}\tabularnewline
\hline 
\hline 
\multirow{2}{*}{Glass} & Water & 63\ensuremath{\pm}5\textdegree{} & 20\ensuremath{\pm}2\textdegree{}\tabularnewline
\cline{2-4} \cline{3-4} \cline{4-4} 
 & n-Hexadecane & <10\textdegree{} & <10\textdegree{}\tabularnewline
\hline 
\multirow{2}{*}{PFOTS} & Water & 122\ensuremath{\pm}1\textdegree{} & 93\ensuremath{\pm}2\textdegree{}\tabularnewline
\cline{2-4} \cline{3-4} \cline{4-4} 
 & n-Hexadecane & 88\ensuremath{\pm}2\textdegree{} & 56\ensuremath{\pm}5\textdegree{}\tabularnewline
\hline 
\end{tabular}\caption{Dynamic contact angles measurements \label{tab:Contact-Angles}}
\end{table}

\subsubsection{Adhesion test}

Adhesion measurements were performed on a custom force measurement
setup developed in-house (Figure \ref{fig:Setup}). A fiber optic
displacement sensor (\emph{Philtec D20, PHILTEC, Inc. USA}) together
with a steel bending beam (spring constant = 68.1 N/m) constitutes
the vertical force sensor. Beam deflection was calibrated using 4
different known weights to get the corresponding force. A plastic
3D printed substrate holder was glued to the end of the bending beam.
The holder was designed to enable switching from one substrate to
another without removing any glue. It also had transparent side walls
which allowed us to fill it with water for the underwater experiments
as well as observe the contact. The sensor was mounted on a stage
consisting of a X-piezo, used for precise lateral movements, XYZ motors,
for coarse movements and a goniometer, for adjusting substrate alignment
with the optics. Additionally, a separate Z-piezo, fixed upright,
was used for vertical up-down motion, bringing the insect in contact
with the substrate from the top. A 3 axis manual micro-manipulator
together with the rotatable steel ball allowed good alignment of the
beetle's foot to the substrate. A coaxial illuminated tube microscope
(\emph{Navitar}) with 2$\times$ objective and a stereo-microscope
with 1$\times$ objective (\emph{Wild Heerbrugg}) fit with cameras
were used to record the sample contact with the substrate from ventral
and side views respectively. Pad contact area was visualized through
the substrate in total reflection mode by the ventral view camera.
The data acquisition from the force sensor and cameras, together with
the appropriate piezo motion steps were synchronized using a custom
LABVIEW program. Force data was acquired at a sample rate of 984 Hz,
averaged to 512 points per motion step. Videos were recorded at 20
frames per second.

The vertical and lateral piezos were together used to perform approach-retract
adhesion tests with the substrate to get the pull-off force. However,
instead of a simple down-up motion, some additional motion steps were
included (Figure \ref{fig:Setup}). A 100 \textgreek{m}m lateral sliding
motion was done after the leg makes contact, to give it a proximal
pull which will orient its hairs with the substrate \cite{RN19}.
An additional 10 \textgreek{m}m compression step (approach) was done
to ensure all hairs were loaded in compression and made good contact
with the surface. A short pause (1 s) was introduced to minimize any
viscoelastic effects before finally retracting the leg away from the
substrate. All approach, retract and lateral slide motion was done
at a speed of 62.5 \textgreek{m}m/s. Ventral view video recordings
were used for contact area extraction while the side view imaging
was used to visually aid orienting the pad with the substrate before
a test.

For underwater experiments, 1 ml Milli-Q water was pipetted into the
substrate holder. In order to achieve an underwater contact without
a trapped air bubble, the water was first degassed in a vacuum chamber
at 10 mbar pressure for 3 hours and used immediately. Before experiment,
the pad was repeatedly brought into contact with the surface 10 times
to equilibriate the pad system. 5 force measurements were subsequently
performed on a fresh spot of the substrate and averaged for data analysis.
Experiments were repeated with 5 individual beetles for each type
of contact and substrate. In total, 30 beetles were tested.

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.5]{Figure1-Setup_schematic}
\par\end{centering}
\caption{\label{fig:Setup} Adhesion test setup. The recorded force data and
contact area of distal pad are shown in the plot, in which, the shaded
regions represent distinct piezo motion steps, negative force values
represent attraction and the minimum force peak during the final retraction
step is the adhesion force.}
\end{figure}

\subsubsection{Field desorption mass spectroscopy}

A previous study on archanids \cite{RN223} reported that its secreted
fluid does not dissolve in water. To confirm if even the beetle retains
its adhesive fluid underwater, we perform Field Desorption Mass Spectroscopy
(FDMS) using a ZAB 2-SE-FPD spectrometer (VG Instruments). The measurements
were done on the extracted secretions before and after rinsing its
legs with water. The middle leg of an Asian ladybird (\emph{Harmonia
axyridis}) was immersed in 50 \textmu L THF for 20 min and then transferred
to the measurement chamber of the FDMS. As a reference, pure THF was
used. The second middle leg of the same ladybird was subsequently
immersed in 100 \textmu L Milli-Q water for 15 min, then in THF for
20 min and then transferred to the measurement chamber of the spectrometer.
Molecular composition was extracted from the peak positions of the
FDMS spectra.

\subsubsection{Data analysis}

Extraction of pull-off force from force data, image processing, plotting
and statistical analysis were all performed in ``\emph{Buggee}'',
a tool written in Python using open-source libraries for synchronous
analysis of force data and video recordings (\url{https://github.com/PranavSudersan/Buggee}).

For measurements in air, the pull-off force was defined as the minimum
negative force during the final retraction step (Figure \ref{fig:Setup}).
For underwater measurements, an additional correction was necessary.
When the beetle was submerged underwater, its contact line at the
water surface shifted, which influenced the force readout due to surface
tension effects. This effect needed to be canceled. So, a ``background''
force data was recorded, where the submerged beetle makes no contact
with the substrate. This background data was then subtracted from
a typical force data with substrate contact to correct for the external
surface tension effects. The pull-off force was subsequently calculated
from the minima as before.

Data sets were compared for statistical differences using pairwise
Student t-test and their corresponding p-value and Common Language
Effect Size (CLES) are reported. Shapiro-Wilk test was done for each
data set to verify a normal distribution of its residuals and Levene's
test was done to check for variance homogeneity, to validate the t-test
assumptions. Bonferroni's correction was used to account for multiple
comparison between groups.

\subsection{Results}

\subsubsection{Adhesion measurements}

In air, adhesion forces of the distal pad of ladybug beetle against
glass and PFOTS are rather similar; no significant differences are
detected (Figure \ref{fig:Effect-of-contact} and Table \ref{tab:Statistical-analysis}).
In contrast, underwater adhesion on PFOTS surface is significantly
larger than on glass (p < 0.001). This stronger adhesion on PFOTS
is observed both in the presence and absence of a trapped bubble.
In both cases, adhesion force reach similar values as in air. In contrast,
on glass, adhesion underwater is significantly reduced as compared
to dry conditions, irrespective of the presence of a trapped bubble
(p $\leq$ 0.002). Underwater adhesion on glass tends to be slightly
higher in the presence of a bubble (CLES = 0.84). However, this difference
is small and not statistically significant (p = 0.07).

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{Figure2-Expt_effect_of_contact}\caption{\label{fig:Effect-of-contact}Box-and-whisker plot showing single
leg adhesion force measurements of ladybug beetle (\emph{Coccinella
septempuctata}) on glass and PFOTS substrates in air and underwater
conditions. The two types of contact during underwater experiments
are represented separately: \textquotedblleft\emph{bubble}\textquotedblright{}
and \textquotedblleft\emph{no bubble}\textquotedblright . Crosses
represent theoretical predictions of adhesion force, while, circles
represent the contribution of the bubble itself, calculated from the
capillary bridge model. In the model, hair diameter = 4 \textgreek{m}m,
pad diameter = 200 \textgreek{m}m, hair height = 40 \textgreek{m}m,
N\protect\textsubscript{hairs} = 500, V\protect\textsubscript{fluid}
= 4.2 fL and V\protect\textsubscript{bubble} = 1.2 nL. Interfacial
tension of the adhesive fluid in air and water were assumed to be
24 mN/m and 48 mN/m respectively and water surface tension is 72 mN/m.}
\end{figure}

Apart from the three depicted contact types, we observed an additional
fourth type which occurred in roughly 25\% of our underwater experiments.
In this scenario, the ventral view recordings show that the hairs
did not appear to contact well with the substrate, unlike the other
three contact types. This ``bad contact'' scenario shows no adhesion
with either glass or PFOTS substrate. While it was not completely
clear why such a contact occurs, there can be two possible reasons.
First, the hairs could get bundled due to a small air meniscus within
the hairs, resulting in their disorientation. Second, a thin water
layer at the substrate interface might not be drained out to allow
the hairs to make contact with the substrate, resulting in a loss
of adhesion.

\subsubsection{Field desorption mass spectroscopy}

FDMS results confirm the presence of adhesive fluid on an Asian ladybird's
leg when underwater (Table \ref{tab:Molecular-distribution-of}).
Molecular weights of the secreted fluid mixtures extracted from the
beetle's leg, without and after immersion in water, show that, except
for two molecular weights (406.8 g/mol and 331.6 g/mol), the chemical
fingerprint remained unchanged This indicates that the adhesion fluid
was not washed away underwater. Probable compounds in the fluid, corresponding
to the resultant molecular weights, include mostly aliphatic hydrocarbons
with traces of aldehydes.

\begin{table}[H]
\centering{}%
\begin{tabular}{|c|c|c|}
\hline 
Without rinsing (g/mol) & After rinsing (g/mol) & Probable compounds\tabularnewline
\hline 
\hline 
324.5 & 324.5 & C\textsubscript{23}H\textsubscript{48}, C\textsubscript{22}H\textsubscript{44}O\tabularnewline
\hline 
 & 331.6 & C\textsubscript{24}H\textsubscript{44}\tabularnewline
\hline 
350.5 & 350.5 & C\textsubscript{25}H\textsubscript{50}\tabularnewline
\hline 
352.5 & 352.5 & C\textsubscript{25}H\textsubscript{52}, C\textsubscript{24}H\textsubscript{48}O\tabularnewline
\hline 
378.5 & 378.5 & C\textsubscript{27}H\textsubscript{54}\tabularnewline
\hline 
404.6 & 404.5 & C\textsubscript{29}H\textsubscript{56}\tabularnewline
\hline 
406.8 &  & C\textsubscript{29}H\textsubscript{58}\tabularnewline
\hline 
432.8 & 432.7 & C\textsubscript{31}H\textsubscript{60}\tabularnewline
\hline 
\end{tabular}\caption{Molecular weights of adhesive fluid secretion of \emph{Harmonia axyridis}
with and without rinsing the beetle's leg in water. Molecular weights
were extracted from the peaks in the FDMS spectra. \label{tab:Molecular-distribution-of}}
\end{table}

\section{Theory}

\subsection{Capillary Bridge Model}

\begin{figure}[H]
\includegraphics[width=0.7\linewidth]{Figure3-Model_schematic}\caption{\label{fig:Model}The capillary bridge model. The hairs make contact
with the substrate in three ways: a) \emph{In air}, where the adhesive
fluid bridges are surrounded by air; b) \emph{Underwater: no bubble},
where the adhesive fluid bridges are fully surrounded by water; c)\emph{
Underwater: bubble}, where part of the adhesive fluid bridges are
inside the bubble while others are outside in water. The corresponding
ventral view contact images of the beetle's pad are shown on the right.}
\end{figure}
We model the hairy pad as an array of $N$ cylindrical rods of length,
$L$, and diameter, $D_{h}$, fixed to a flat circular pad of diameter,
$D_{p}$ (Figure \ref{fig:Model}). The hairs and the pad are assumed
to be perfectly rigid, for simplicity. The tip of each hair has an
adhesive fluid of volume, $V_{f}$, making contact with the substrate.
The fluid is pinned to the circumference of the hair and forms a capillary
bridge of height, $d$. Similar to our experiments, we consider three
types of contacts for the pad: 1) \emph{In air}, 2) \emph{Underwater:
no bubble} and 3) \emph{Underwater: bubble}. In the third case, a
bubble of volume, $V_{b}$, is trapped between the hairs and pinned
to the pad circumference.

To characterize the adhesive fluid and bubble volume we define two
radii, $s_{f}$ and $s_{b}$, respectively, by $V_{f}=\frac{4}{3}\pi s_{f}^{3}$
and $V_{b}=\frac{4}{3}\pi s_{b}^{3}$. Here, $s_{f}$ and $s_{b}$
are the radii of spheres with equivalent volumes. Fluid and bubble
radii are assumed to scale proportional to their corresponding pinned
contact diameter. We thus defined the size parameters, $\phi_{f}=D_{h}/\left(2s_{f}\right)$
and $\phi_{b}=D_{p}/\left(2s_{b}\right)$ for the fluid and bubble
respectively, to conveniently scale their volumes relative to the
hair and pad diameters they were pinned to.

The net force for cases 1 and 2 can be calculated as:

\begin{equation}
F_{net}=Nf\label{eq:f_air/water}
\end{equation}

Here, $f$ is the capillary force of a single fluid bridge at a distance,
$d$, in air ($f_{air}$) or underwater ($f_{water}$).

For case 3, the net force is given by:

\begin{equation}
F_{net}=N_{in}f_{air}+N_{out}f_{water}+f_{bubble}\label{eq:f_bubble}
\end{equation}

Here, $N_{in}$ and $N_{out}$ are the number of hairs inside and
outside the bubble, respectively, $f_{air}$ and $f_{water}$ are
the capillary forces of the fluid bridge inside and outside the bubble,
respectively, and $f_{bubble}$ is the capillary force contribution
due to the bubble meniscus alone at distance $d+L$.

The capillary force, $f$, is the sum of two contributions: surface
tension and Laplace pressure (equation \ref{eq:f_bridge}). Force
versus distance for a single capillary bridge is calculated by Surface
Evolver simulations\cite{RN206}, (described in \ref{subsec:Simulation-Method})
and used to obtain $F_{net}$ as a function of $d$ for each type
of contact. The adhesion force of the complete hairy pad system is
then obtained from the minima of $F_{net}$, where negative force
values represent attraction.

We have considered $f_{air}$ and $f_{water}$ to be distinct terms
because the capillary force by the adhesive fluid will be different
in air and underwater due to its different contact angle and interfacial
tension in each case. Using the Young-Dupr\textcyr{\`\cyre} equations
for each case of fluid-air, fluid-water and water-air interface, one
can derive the following relation for the contact angle of the adhesive
fluid underwater:

\begin{equation}
\cos\theta_{fw}=\frac{\gamma_{fa}\cos\theta_{fa}-\gamma_{wa}\cos\theta_{wa}}{\gamma_{fw}}\label{eq:theta_fw}
\end{equation}

Here, $\theta_{fw}$ and $\theta_{fa}$ are the contact angles of
the adhesive fluid with the substrate in water and air respectively,
$\theta_{wa}$ is the contact angle of water with the substrate in
air, $\gamma_{fa}$ is the surface tension of the adhesive fluid,
$\gamma_{wa}$ is the surface tension of water and $\gamma_{fw}$
is the interfacial tension of the adhesive fluid with water.

All lengths are normalized w.r.t. $s_{f}$ and forces are normalized
w.r.t. $\gamma_{fa}s_{f}$. Interfacial tension values are fixed relative
to $\gamma_{fa}$. Non dimensional bubble volume is expressed as,
$\hat{V}_{b}=V_{b}/s_{f}^{3}$

Geometric parameters and interfacial properties are kept fixed for
all model calculations (Table \ref{tab:Model-parameters}). Here,
we assume the adhesive fluid to be an oil-lke substance and thus the
interfacial tension ratios $\gamma_{wa}/\gamma_{fa}$ and $\gamma_{fw}/\gamma_{fa}$
are taken corresponding to typical values of oil and water. We consider
representative hydrophilic and hydrophobic substrates with $\theta_{fa}$
and $\theta_{wa}$ values corresponding to a typical glass and fluorinated
surface, respectively. Area fraction of the hairs relative to the
pad, $\alpha=ND_{h}^{2}/D_{p}^{2}$, hair aspect ratio, $L/D_{h}$,
and fluid size parameter, $\phi_{f}$, are fixed to values typical
for a ladybug's hairy pad.

First, we calculate force-distance curves for a single pinned liquid
capillary bridge. Second, the effect of substrate on the force-distance
curves of the hairy pad system is compared for each type of contact.
Third, the effect of varying hair diameter, $D_{h}$, on net adhesion
is studied. Finally, we look at the effect of changing the bubble
volume, $\hat{V}_{b}$, on the net underwater adhesion.

\begin{table}[H]
\centering{}%
\begin{tabular}{|l|c|}
\hline 
Property & Value\tabularnewline
\hline 
\hline 
Area fraction, $\alpha$ & 0.1\tabularnewline
\hline 
Hair aspect ratio, $L/D_{h}$ & 10\tabularnewline
\hline 
Water surface tension ratio, $\gamma_{wa}/\gamma_{fa}$ & 3\tabularnewline
\hline 
Fluid-water interfacial tension ratio, $\gamma_{fw}/\gamma_{fa}$ & 2\tabularnewline
\hline 
Fluid size parameter, $\phi_{f}$ & 2\tabularnewline
\hline 
Hydrophilic substrate & $\theta_{fa}=6\lyxmathsym{\textdegree}$, $\theta_{wa}=24\lyxmathsym{\textdegree}$\tabularnewline
\hline 
Hydrophobic substrate & $\theta_{fa}=50\lyxmathsym{\textdegree}$, $\theta_{wa}=120\lyxmathsym{\textdegree}$\tabularnewline
\hline 
\end{tabular}\caption{\label{tab:Model-parameters}Fixed parameters of the capillary bridge
model}
\end{table}

\subsection{Capillary force of a pinned liquid bridge\label{subsec:Capillary-force-of}}

Forces due to a single pinned capillary liquid bridge in contact with
a substrate are obtained via Surface Evolver simulations (Figure \ref{fig:Single-bridge}).
We see that, generally, the shape of the liquid meniscus determines
the strength of its adhesion force. High adhesion is seen for contact
angles less than $\sim$ 70\textdegree{} due to a net negative curvature
of the meniscus, while low adhesion is seen for contact angles greater
than $\sim$ 150\textdegree{} due to its net curvature being close
to zero. The Laplace pressure contribution to the net adhesion force
dominates for contact angles less than 100\textdegree{} (Figure \ref{fig:Single-bridge}b).
Interestingly, its contribution to the adhesion force is mostly non-repulsive
for contact angles greater than 90\textdegree . This is because, the
low volume of the liquid and its pinned contact line prevents the
meniscus from having a high positive curvature due to geometric constraints.
Only for a contact angle of 150\textdegree , the liquid's curvature
becomes positive, manifested in its slightly repulsive Laplace contribution.
Surface tension makes a significant contribution to the net force
only for a small range of contact angles close to 90\textdegree .
For contact angles greater than 150\textdegree , the net adhesion
force approaches zero.

\begin{figure}[H]
\begin{minipage}[t]{0.5\columnwidth}%
\includegraphics[scale=0.5]{Figure4a-Single_bridge_fd}

\subcaption{Force-distance curves}%
\end{minipage}\hfill{}%
\begin{minipage}[t]{0.5\columnwidth}%
\includegraphics[scale=0.5]{Figure4b-Single_bridge_force_contribution}

\subcaption{Force contributions}%
\end{minipage}\caption{\label{fig:Single-bridge}Normalized capillary force of a single liquid
bridge in contact with a substrate and pinned to a circular perimeter
on top. Fluid size parameter, $\phi_{f}=2$. Negative force values
represents attraction. a) Force-distance curves are shown for different
contact angles of the liquid with the substrate. b) Adhesion forces,
calculated from the minima of the corresponding force-distance curves,
are plotted as a function of contact angle with the substrate, together
with its Laplace and surface tension components (equation \ref{eq:f_bridge}).
Simulation snapshots of the liquid meniscus corresponding to angles
6\textdegree{} and 150\textdegree{} are depicted.}
\end{figure}

The force-distance curves show a general trend of being repulsive
at small distances (Figure \ref{fig:Single-bridge}a). This is a result
of the pinned contact line on the top. A limited volume is available
for the liquid to occupy when the gap distance is small, causing the
meniscus shape to bulge outwards near the pinned contact line. This
creates a net positive curvature, resulting in a positive Laplace
pressure and thus repulsion. Without pinning, the capillary forces
would have shown high attractive forces on a hydrophilic substrate\cite{RN93}.

\subsection{Effect of the substrate\label{subsec:Capillary-Bridge-Model:}}

The normalized force-distance curves of a hairy pad system on a hydrophilic
and hydrophobic substrate are predicted based on the capillary bridge
model and compared for the different contact types (Figure \ref{fig:Effect-of-substrate}).
The forces in each case are calculated from equations (\ref{eq:f_air/water})
and (\ref{eq:f_bubble}) for fixed geometric and interfacial properties
(Table \ref{tab:Model-parameters}).

On the hydrophilic substrate ($\theta_{wa}=24\lyxmathsym{\textdegree}$),
highest adhesion is seen when the hairs contact in air, while lowest
adhesion occurs underwater without a trapped bubble. The presence
of a bubble leads to intermediate force values. In contrast, on a
hydrophobic substrate ($\theta_{wa}=120\lyxmathsym{\textdegree}$),
highest adhesion is seen for the underwater case without a trapped
bubble, much larger than in air. When a bubble is present, the forces
are only slightly larger than in air.

\begin{figure}[H]
\centering{}\includegraphics{Figure5-Model_effect_of_substrate}\caption{\label{fig:Effect-of-substrate}Theoretical force-distance curves
of a hairy pad system on a hydrophilic and hydrophobic substrate in
air and underwater conditions. A negative force value represents attraction.
Normalized forces are calculated from the capillary bridge model,
with model parameters listed in Table \ref{tab:Model-parameters}.
The bubble's contribution to the net force for an \emph{underwater:
bubble} contact is denoted by plus symbols. Insets represent the \emph{underwater:
bubble }contact for each substrate.}
\end{figure}

The observed trend in forces can be explained by how the adhesive
fluid wets the surface in each case. On a hydrophilic substrate, the
contact angle of the oily adhesive fluid is 6\textdegree , when surrounded
by air (Table \ref{tab:Model-parameters}) and 150\textdegree , when
surrounded by water (equation (\ref{eq:theta_fw})). This results
in the meniscus shape to have a net negative and slightly positive
curvatures, respectively, resulting in strong adhesion in air and
poor adhesion underwater. On a hydrophobic substrate however, the
contact angles of the fluid in air and water are 50\textdegree{} and
1\textdegree , respectively. In both cases, the contact angles are
low, resulting in strong adhesion in both media. Additionally, the
interfacial tension of the oily fluid underwater ($\gamma_{fw}$)
is twice that of in air ($\gamma_{fa}$). Thus, we see a higher capillary
adhesion for the \emph{underwater: no bubble} case when compared to
\emph{in air} (Figure \ref{fig:Oil-contact-images}). Note that the
contact area is fixed in all cases by keeping the area fraction and
$D_{p}/D_{h}$ constant. Thus the observed effects is not a result
of changing contact area, but rather on the nature of capillary forces.

\begin{figure}[H]
\includegraphics{Figure-8-contact_angle_schematic}\caption{\label{fig:Oil-contact-images}Simulation snapshots of oil capillary
meniscus in contact with glass and PFOTS in air and underwater conditions.
The corresponding interfacial tension, \textgreek{g}, and contact
angle, \ensuremath{\theta}, used to predict the ladybug's adhesion
are labeled for each case.}
\end{figure}

The net force in the \emph{underwater: bubble} case mainly depends
on the proportion of hairs inside and outside the bubble (equation
(\ref{eq:f_bubble})). For the given bubble volume, only part of the
hairs are inside the bubble for the hydrophilic substrate, while,
all the hairs are inside the bubble for the hydrophobic substrate.
Therefore, the force curve lies between \emph{in air} and \emph{underwater:
no bubble} cases for a hydrophilic substrate, and closely follows
the \emph{in air} case for a hydrophobic substrate.

We see that the bubble itself doesn't contribute much to the net force
on either substrate (Figure \ref{fig:Effect-of-substrate}). Its contribution
even is slightly repulsive on the hydrophilic substrate due to the
positive curvature of the bubble, and slightly attractive on the hydrophobic
substrate due to its negative curvature. This small contribution is
manifested by the slightly higher adhesion for \emph{underwater: bubble}
relative to \emph{in air} for the hydrophobic substrate, since all
hairs are within the bubble in this case.

\subsection{Effect of bubble volume}

\begin{figure}[H]
\includegraphics{Figure7-Model_effect_of_bubble_volume}\caption{\label{fig:Effect-of-bubble}Normalized adhesion force of a hairy
pad system as a function of bubble volume, $\hat{V}_{b}$, for the
\emph{underwater: bubble} contact type. Adhesion forces are calculated
from the minima of the respective force-distance curves. Negative
force value represents attraction. Pad to hair diameter ratio ($D_{p}/D_{h}$)
is kept fixed. Highlighted regions represent entrapment of all hairs
within the bubble.}
\end{figure}
The volume of the bubble can influence its capillary force contribution,
as well as change the relative proportion of hairs inside and outside
it. To investigate this, we varied the bubble volume, $\hat{V}_{b}$,
and compared the maximum adhesion force on both hydrophilic and hydrophobic
substrates (Figure \ref{fig:Effect-of-bubble}). The contribution
of the bubble to the net adhesion force is small regardless of its
volume, when compared to the whole pad. Further, opposite trends of
adhesion is seen on the two substrates with changing $\hat{V}_{b}$.

From previous section, we know that on the hydrophilic substrate,
fluid bridges outside the bubble show poor adhesion due to the positive
curvature of their meniscus. Thus, decreasing $\hat{V}_{b}$ decreases
the adhesion force due to a larger proportion of hairs being outside
the bubble. In contrast, on the hydrophobic substrate, fluid bridges
outside the bubble show higher capillary forces, due to its low contact
angle and high interfacial tension in water. Thus, adhesion force
increases for a hydrophobic substrate as the bubble size decreases.

A smaller $\hat{V}_{b}$ results in increased, but small, attraction
by the bubble on both types of substrates. For larger values of $\hat{V}_{b}$
however, the force trend for the whole pad mostly follows that of
the bubble. This is because the bubble gets big enough to entrap all
the hairs inside it. Thus, the force contribution due to the fluid
bridges remain unchanged, and only the bubble's contribution drives
the slight variation in the pad's adhesion at high $\hat{V}_{b}$.
Once the bubble is small enough such that part of the fluid bridges
start making contact in water, the force trend changes, with a steep
decrease (increase) in adhesion force on hydrophilic (hydrophobic)
substrate as the volume decreases.

\subsection{Effect of hair diameter }

\begin{figure}[H]
\includegraphics[scale=0.5]{Figure6-Model_effect_of_hair_size}\caption{\label{fig:Effect-of-hair}Normalized adhesion force of a hairy pad
system on a hydrophilic and hydrophobic substrate as a function of
hair diameter, $D_{h}$. Volume of each fluid bridge, $V_{f}$, scales
relative to $D_{h}$ based on the parameter $\phi_{f}=2$. Adhesion
forces are calculated from the minima of the respective force-distance
curves, based on the capillary bridge model. A negative value represents
attraction. The bubble's contribution to the net force for an \emph{underwater:
bubble} contact is denoted by plus symbols. Pad diameter and bubble
volume are kept fixed. All lengths are scaled relative to $D_{p}$.}
\end{figure}
Hairs on a ladybug's pad terminate in various shapes like ``discoidal''
or ``pointed''. We studied this geometric effect on adhesion by
changing the hair diameter, $D_{h}$ (Figure \ref{fig:Effect-of-hair}).
Here, the pad diameter, total hair contact area and bubble volume
are constant since $D_{p}$, $\alpha$ and $\hat{V}_{b}$ are kept
fixed. The adhesive fluid volume is again assumed to scale relative
to the hair diameter ($\phi_{f}=2$).

Adhesion force increases with decreasing $D_{h}$ for both hydrophilic
and hydrophobic substrates in all contact types. This is consistent
with the ``contact splitting'' theory, which predicts higher adhesion
when the contact is split into many small contact points\cite{RN24}.
Reducing the hair diameter results in two competing effects: 1) capillary
force due to a single fluid bridge decreases due to its smaller size
and ``self-similar'' scaling assumption ($f\sim D_{h}$), which
decreases the net force, and 2) total number of fluid bridges increases
since the total hair contact area is assumed to be fixed ($N\sim1/D_{h}^{2}$),
which increases the net force. The second effect dominates, resulting
in a higher adhesion force as $D_{h}$ decreases.

Similar to the trend in Figure \ref{fig:Effect-of-substrate}, contact
\emph{in air} shows the highest adhesion force on a hydrophilic substrate
for the given range of hair diameters, while on a hydrophobic substrate,
\emph{underwater: no bubble} shows highest adhesion. \emph{Underwater:
bubble} contact shows intermediate adhesion between \emph{in air}
and \emph{underwater: no bubble} contact types.

The bubble's contribution gets repulsive as hair diameter decreases
for both substrates (Figure \ref{fig:Effect-of-hair}). Since the
aspect ratio $L/D_{h}$ is fixed (Table \ref{tab:Model-parameters}),
decreasing the hair diameter also decreases its length. Since the
bubble's volume is kept constant, it will then have a lesser space
available to occupy between the pad and the substrate, This results
in it bulging outwards near the pinned contact line on the top, causing
repulsion.

\section{Discussion}

Our experiments demonstrate that the ladybug beetle can attach underwater
to a hydrophobic substrate even without a bubble trapped around its
hairs. A previous study\cite{RN87} had hypothesized that a bubble
is necessary for underwater attachment in beetles. This is, however,
only true for hydrophilic substrates, where a trapped bubble can facilitate
underwater adhesion due to the hairs making contact in a dewetted
environment. For a hydrophobic substrate, the adhesion is similar
regardless of whether the contact occurs in air or underwater conditions,
with or without a trapped bubble. Our theoretical calculations further
show that the bubble by itself has a negligible capillary contribution
to the net underwater adhesion of the pad. Direct force measurement
of a single similarly sized bubble making contact with a hydrophobic
substrate shows a maximum adhesion less than 50 \textgreek{m}N, which
further validates that the bubble's contribution is insignificant
(\ref{subsec:Capillary-force-due}).

Predictions of the ladybug's adhesion from the capillary bridge model
agree with our experimental results (Figure \ref{fig:Effect-of-contact}).
In underwater conditions without a trapped bubble, adhesion on a hydrophobic
substrate is significantly larger than on a hydrophilic substrate.
This is explained by the different oil interfacial tension and contact
angles with the substrates in air and underwater, which determines
the capillary adhesive force in each case (Figure \ref{fig:Oil-contact-images}).
However, the experiments don't show the predicted $\sim$ 2.6 times
increase in underwater adhesion relative to that in air on the hydrophobic
PFOTS surface. This discrepancy could be due to our assumptions of
the oily fluid's interfacial properties. If we choose $\gamma_{fa}$=30
mN/m and $\gamma_{fw}$=40 mN/m, the corresponding increase in adhesion
will be $\sim$ 1.7, closer to our experimental value of $\sim$ 1.
The resulting change in $\theta_{fa}$ and $\theta_{fw}$ will further
decrease this number. Direct measurement of the fluid's interfacial
properties is thus essential to better predict the insect's adhesion,
and will be a subject of future study. Further, due to surface inhomogeneities,
not all the hairs might be able to completely drain the interfacial
water layer, in order for the adhesive fluid to make a direct contact
with the substrate. This can further reduce underwater adhesion, in
comparison to our theoretical predictions which assumes a perfect
contact.

In the model, we assume that all the hairs detach simultaneously to
give a theoretical maximum achievable adhesion force. In our experiments,
however, the pad always makes contact with the substrate at a random
orientation, which is difficult to control precisely. During detachment,
the pad typically peels off from its proximal to distal end rather
than detach simultaneously. Our model also assumes the hairs to be
of similar geometry, unlike the beetle's pad which has a distribution
of flat or pointed tipped hairs. Thus, it's not surprising that the
model overestimates the adhesion forces. The predictions are however
in the same order of magnitude as experiments, and the qualitative
trend is consistent for both hydrophilic and hydrophobic substrates
in air and underwater.

Our study provides further validation that capillary forces by the
adhesive fluid control the insect's adhesion and van der Waals contribution,
if any, must be negligible. Further, the capillary forces can even
enable insect attachment underwater depending on the substrate chemistry.
When underwater, without a trapped bubble, the pad adheres strongly
to a hydrophobic substrate, but poorly to a hydrophilic substrate,
even though it shows similarly strong adhesion to both substrates
in air. This behavior can only be explained by capillary forces. Our
preliminary FDMS results further provides validation to our assumption
that the adhesive fluid can form capillary bridges with the substrate
in water, instead of getting washed away (Table \ref{tab:Molecular-distribution-of}).

The findings can also be extended to other animals relying on oily
secretions for adhesion. Ants, for example, show similar adhesion
on hydrophobic substrates under wet and dry conditions\cite{RN213},
similar to what we see in a ladybug. Recent adhesion experiments on
geckos reveal that they can attach well to fluoropolymer substrates
underwater while they show little adhesion to the same substrate in
air\cite{RN199,RN15}. Geckos are thought to rely on van der Waals
forces via dry contact with the substrate \cite{RN202}, although
recent observations of phospholipid footprints left behind walking
geckos \cite{RN205} could change that picture. Since geckos adhere
poorly to PTFE (surface energy \textasciitilde{} 20 mN/m) one can
speculate that the phospholipid material has a higher surface energy,
and consequently makes a higher contact angle with PTFE in air. Let
us assume the phosopholipid substance to be a fluid similar to oil
with $\gamma_{fa}$= 30 mN/m and $\gamma_{fw}$= 42 mN/m such that
its contact angle with PTFE is 80\textdegree . Equation \ref{eq:theta_fw}
then gives us an underwater contact angle of 70\textdegree{} for the
phospholipid fluid. Thus, on a PTFE surface, the capillary bridge
model can predict a higher adhesion underwater than in air due to
its lower contact angle and higher interfacial energy underwater.
Based on similar assumptions, we predict the net adhesion force for
the gecko on different substrates (Figure \ref{fig:Comparison-with-model}).
The adhesion force predictions are in good qualitative agreement with
the whole animal experimental shear force values reported for the
gecko, with the trend of higher adhesion in air than underwater for
glass, similar adhesion in air and underwater for PMMA/OTS-SAM and
lower adhesion in air than underwater for PTFE. We, thus, propose
that the underwater experiments performed on geckos \cite{RN199,RN15}
indicate a capillary contribution to gecko adhesion. We suggest performing
single seta adhesion force tests similar to Autumn et. al. \cite{RN202}
using a hydrophilic and fluorinated probe in air and underwater conditions
to validate the role of capillary contributions to gecko adhesion.

\begin{figure}[H]
\includegraphics{Figure-9-Gecko_comparison}\caption{\label{fig:Comparison-with-model}Whole animal adhesion force of geckos
on various substrates. Experimental shear adhesion values are reproduced
from Stark et. al. \cite{RN15}. Normal adhesion forces for each gecko
toe are theoretically estimated from the capillary bridge model, with
hair diameter = 400 nm, toe diameter = 4 mm, adhesive fluid volume
= 4.19x10\protect\textsuperscript{-3} fL and 10\% hair coverage.
\textquotedblleft\emph{Underwater: no bubble}\textquotedblright{}
type contact is assumed for the \textquotedblleft Wet\textquotedblright{}
case. Net adhesion force is calculated by assuming 5 toes on each
leg and 4 legs in total on a gecko. Interfacial tension of the phospholipid
layer (PL) in air and water are assumed to be 30 mN/m and 42 mN/m
respectively. PL contact angles with glass, PMMA, OTS-SAM and PTFE
are assumed to be 6\textdegree , 10\textdegree , 20\textdegree{} and
80\textdegree{} respectively. The corresponding water contact angles
are 50\textdegree , 85\textdegree , 94\textdegree{} and 97\textdegree{}
respectively, as reported in Stark et. al. \cite{RN15}.}
\end{figure}

We have so far limited our analysis to only smooth substrates. Insects
in the real world, however, interact with rough substrates very often.
Previous studies \cite{RN136} have shown that substrate roughness
is a more dominant parameter than substrate chemistry in controlling
insect adhesion. Future work will explore how roughness can impact
the net capillary force. Our study can have potential applications
in the design of bio-inspired materials to achieve underwater adhesion
via capillary bridges. Bubble can possibly be used to control underwater
adhesion by changing the relative proportion of the arrays inside
and outside the bubble. A suitable choice of adhesive fluid can be
made tailored to the substrate and environment of application to achieve
optimal adhesion performance.

\section{Conclusions}

Ladybug beetles rely primarily on their oily fluid secretion to adhere
to surfaces in both air and underwater conditions. The beetle can
attach underwater on a hydrophobic substrate even without a trapped
air bubble within its hairy pad, although it loses this ability on
a hydrophilic substrate. This is explained theoretically by the different
contact angle and interfacial tension of the adhesive fluid in air
and underwater conditions. Further, the bubble itself has a negligible
capillary contribution to the total force. The trapped bubble can
promote adhesion only on a hydrophilic substrate by providing an air
medium to the adhesive fluid bridges inside it. Oil wettability, thus,
primarily controls the insect's adhesion in any given condition. A
similar argument also explains previously reported underwater measurements
in geckos \cite{RN15}, which highlights the possibility of capillary
contributions to gecko adhesion. Future studies should characterize
the fluid secretion's interfacial properties with a particular substrate
to better understand the nature of the animal's adhesion.

\section{Acknowledgement}

We are grateful to Eduard Arzt and Ren\textcyr{\`\cyre} Hensel for
fruitful discussions. This work was funded by \emph{Deutsche Forschungsgemeinschaft}
(Grant number: PI 1351/2-1).

\appendix

\section{Appendix}

\subsection{Simulation method: Single capillary bridge \label{subsec:Simulation-Method}}

\setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}} 
\setcounter{equation}{0} \renewcommand{\theequation}{A.\arabic{equation}} Capillary
force due to a single adhesive fluid or bubble meniscus (termed ``capillary
bridge'') is calculated by performing simulations in Surface Evolver
\cite{RN206}, similar to the method described by De Souza et. al.\cite{RN93}.
A simple cubic geometry, mimicking the capillary bridge, of constant
volume, $V$, is defined as the initial condition with an interfacial
tension, $\gamma$, with the surrounding medium. Interfacial tension
of the capillary bridge with the substrate is given by $\gamma\cos\theta$,
where $\theta$ is the corresponding contact angle inside the bridge.
For the case of a bubble meniscus, $\theta$ is defined w.r.t. the
surrounding water, since $\theta$ can also directly characterize
the substrate wettability. The capillary bridge spans a gap distance
$d$ between the top face and the substrate. The boundary conditions
are set corresponding to a pinned contact line of diameter $D$ on
the top face and constant interfacial tension with the substrate on
the bottom. All lengths are normalized relative to length $s=\left(3V/4\pi\right)^{1/3}$.
An appropriate geometry refinement routine is chosen to evolve the
capillary bridge shape to its minimum energy state. The normalized
total capillary force, $\hat{f}=f/\gamma s$, is the sum of the Laplace
pressure and surface tension contributions , where:

\begin{equation}
f=f_{laplace}+f_{surface\,tension}=\varDelta P_{laplace}A_{_{bottom}}+2\pi R_{bottom}\gamma\sin\theta\label{eq:f_bridge}
\end{equation}

Here, $\varDelta P_{laplace}$ is the Laplace pressure of the equilibrium
capillary bridge, $A_{bottom}$ is the contact area of the capillary
bridge with the substrate at bottom and $R_{bottom}$ is the corresponding
radius of contact, all obtained from the simulation output for the
equilibrium surface.

The gap distance $d$ is varied stepwise and the capillary force is
calculated each time to obtain force-distance curves for a particular
choice of $D$ and $\theta$.

\subsection{Single capillary bridge: Effect of volume}

Surface Evolver simulation results showing the effect of volume on
the maximum capillary force of a single fluid bridge. Since the fluid
is pinned at the top to the same diameter, D, a smaller volume would
result in high interfacial curvatures, which increases the capillary
force due to the negative Laplace pressure. In this case, small contact
angles lead to a greater increase in adhesion.

\begin{figure}[H]

\begin{centering}
\includegraphics{FigureS2-Effect_of_fluid_volume}\caption{Normalized maximum capillary force for a single bridge as a function
of fluid volume}
\par\end{centering}
\end{figure}

\subsection{Capillary Bridge Model: Effect of hair diameter at constant fluid
volume}

Here, instead of scaling the fluid volume relative to the hair diameter,
we now assume a fixed total fluid volume distributed equally among
the $N$ hairs. Total fluid volume, $V_{total}=NV_{f}=2000$. Hair
diameter is varied while keeping the total hair contact area constant.
Length is in arbitrary units. Forces increase at a much smaller rate
on decreasing diameter when compared to the case with self-similar
scaling of fluid volume (Figure \ref{fig:Effect-of-hair}).

\begin{figure}[H]
\includegraphics{FigureS3-Effect_of_hair_size(constant_vol)}

\caption{Normalized adhesion force of hairy pad system on a hydrophilic and
hydrophobic substrate as a function of hair diameter ($D_{h}$), calculated
from the capillary bridge model. The total adhesive fluid volume is
fixed to 2000. Adhesion forces are calculated from minima of the respective
force-distance curves. Negative force value represents attraction.
The bubble's contribution to the net force for an \emph{underwater:
bubble} contact is denoted by plus symbols. Bubble volume and pad
diameter are kept fixed. All lengths are scaled relative to $D_{p}$.}
\end{figure}

\subsection{Statistical comparison}

Pairwise statistical comparison of single leg adhesion force measurements
of the ladybug beetle (\emph{Coccinella septempuctata}) for each contact
type and substrate are shown (Table \ref{tab:Statistical-analysis}).
The uncorrected p-values and Common Language Effect Size (CLES) are
obtained from post-hoc pair-wise Student t-test between A and B while
keeping the third parameter fixed. p-values showing statistically
significant difference between A and B are in boldface. The condition
for statistical significance is based on the Bonferroni-corrected
critical p-value of 0.008.

\begin{table}[H]
\noindent \begin{centering}
\begin{tabular}{|>{\raggedright}m{0.15\linewidth}|>{\raggedright}m{0.15\linewidth}|>{\raggedright}m{0.15\linewidth}|>{\centering}m{0.15\linewidth}|>{\centering}m{0.15\linewidth}|}
\hline 
Fixed & A & B & p-value & CLES\tabularnewline
\hline 
\hline 
In air & PFOTS & Glass & 0.959 & 0.48\tabularnewline
\hline 
Underwater: bubble & PFOTS & Glass & 0.011 & 0.96\tabularnewline
\hline 
Underwater: no bubble & PFOTS & Glass & \textbf{< 0.001} & 1.0\tabularnewline
\hline 
PFOTS & In air & Underwater: bubble & 0.897 & 0.48\tabularnewline
\hline 
PFOTS & In air & Underwater: no bubble & 0.828 & 0.48\tabularnewline
\hline 
PFOTS & Underwater: bubble & Underwater: no bubble & 0.721 & 0.44\tabularnewline
\hline 
Glass & In air & Underwater: bubble & \textbf{0.002} & 1.0\tabularnewline
\hline 
Glass & In air & Underwater: no bubble & \textbf{< 0.001} & 1.0\tabularnewline
\hline 
Glass & Underwater: bubble & Underwater: no bubble & 0.07 & 0.84\tabularnewline
\hline 
\end{tabular}
\par\end{centering}
\caption{\label{tab:Statistical-analysis}}
\end{table}

\subsection{Capillary force due to a bubble\label{subsec:Capillary-force-due}}

Capillary force of a single bubble against a PFOTS surface are compared
for two different volumes. The volumes correspond to the expected
range in the case of the trapped bubble in a ladybug. Here, the bubble
is pinned to a micropatterned PDMS substrate on the top. The maximum
adhesion force of any of the bubble never exceeds 50 \textgreek{m}N,
significantly lower than the beetle's underwater adhesion to the same
substrate (> 400 \textgreek{m}N). Thus, the bubble's contribution
to adhesion in the ``\emph{underwater: bubble}'' contact of a ladybug's
pad should be negligible.

\begin{figure}[H]
\begin{centering}
\includegraphics{FigureS4-Expt_bubble_force}\caption{Capillary force of the bubble}
\par\end{centering}
\end{figure}

\bibliography{references}

\end{document}