### 2021-05-11 source/tex/supplementary.tex
final formatting edits for submission to JEB

\documentclass[english]{achemso}
\setkeys{acs}{keywords = true, email=false}
\usepackage[T2A,LGR,T1]{fontenc}
\usepackage{babel}
\usepackage{array}
\usepackage{textcomp}
\usepackage{url}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subscript}
%\usepackage[unicode=true,pdfusetitle,
% bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
% breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
% {hyperref}
%\usepackage{lineno}
%\linenumbers

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.

\title{Wetting of the tarsal adhesive fluid controls underwater adhesion in ladybug beetles: Supplementary information}

\author{Pranav Sudersan}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\author{Michael Kappl}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\author{Bat-El Pinchasik}

\affiliation{School of Mechanical Engineering, Tel Aviv University, Tel Aviv-Yafo,
Israel}

\author{Hans-J\"{u}rgen Butt}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\author{Thomas Endlein}

\affiliation{Max Planck Institute for Polymer Research, Ackermannweg 10, 55128
Mainz, Germany}

\DeclareRobustCommand{\greektext}{%
  \fontencoding{LGR}\selectfont\def\encodingdefault{LGR}}
\DeclareRobustCommand{\textgreek}[1]{\leavevmode{\greektext #1}}
\ProvideTextCommand{\~}{LGR}[1]{\char126#1}

\DeclareRobustCommand{\cyrtext}{%
  \fontencoding{T2A}\selectfont\def\encodingdefault{T2A}}
\DeclareRobustCommand{\textcyr}[1]{\leavevmode{\cyrtext #1}}

\newcommand{\lyxmathsym}[1]{\ifmmode\begingroup\def\b@ld{bold}
  \text{\ifx\math@version\b@ld\bfseries\fi#1}\endgroup\else#1\fi}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\SectionNumbersOn
%\setkeys{acs}{doi = true}
\setkeys{Gin}{width=\linewidth}
\usepackage{subcaption}
\usepackage[pagewise]{lineno}
\linenumbers

\makeatother

\begin{document}
%reset figure/table/equation numbers and include prefix
\renewcommand{\thesection}{S\arabic{section}}
\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} 
\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} 
\setcounter{equation}{0} \renewcommand{\theequation}{S\arabic{equation}}

\section{Simulation method: Single capillary bridge \label{subsec:Simulation-Method}}

Capillary force due to a single adhesive fluid or bubble meniscus (termed ``capillary
bridge'') was calculated by performing simulations in Surface Evolver
\cite{RN206}, similar to the method described by \citet{RN93}.
A simple cubic geometry, mimicking the capillary bridge, of constant
volume, $V$, was defined as the initial condition with an interfacial
tension, $\gamma$, with the surrounding medium. Interfacial tension
of the capillary bridge with the substrate is given by $\gamma\cos\theta$,
where $\theta$ is the corresponding contact angle inside the bridge.
For the case of a bubble meniscus, $\theta$ is defined w.r.t. the
surrounding water, since $\theta$ can also directly characterise
the substrate wettability. The capillary bridge spans a gap distance
$d$ between the top face and the substrate. The boundary conditions
were set corresponding to a pinned contact line of diameter $D$ on
the top face and constant interfacial tension with the substrate on
the bottom. All lengths were normalised relative to length $s=\left(3V/4\pi\right)^{1/3}$.
An appropriate geometry refinement routine was chosen to evolve the
capillary bridge shape to its minimum energy state. The normalised
total capillary force, $\hat{f}=f/\gamma s$, is the sum of the Laplace
pressure and surface tension contributions , where:

\begin{equation}
f=f_{laplace}+f_{surface\,tension}=\varDelta P_{laplace}A_{_{bottom}}+2\pi R_{bottom}\gamma\sin\theta\label{eq:f_bridge}
\end{equation}

Here, $\varDelta P_{laplace}$ is the Laplace pressure of the equilibrium
capillary bridge, $A_{bottom}$ is the contact area of the capillary
bridge with the substrate at bottom and $R_{bottom}$ is the corresponding
radius of contact, all obtained from the simulation output for the
equilibrium surface.

The gap distance $d$ was varied stepwise and the capillary force was
calculated each time to obtain force-distance curves for a particular
choice of $D$ and $\theta$.

%\section{Single capillary bridge: Effect of volume}
%
%Surface Evolver simulation results showing the effect of volume on
%the maximum capillary force of a single fluid bridge. Since the fluid
%is pinned at the top to the same diameter, D, a smaller volume would
%result in high interfacial curvatures, which increases the capillary
%force due to the negative Laplace pressure. In this case, small contact
%angles lead to a greater increase in adhesion.
%
%\begin{figure}[H]
%
%\begin{centering}
%\includegraphics{FigureS2-Effect_of_fluid_volume}\caption{Normalised maximum capillary force for a single bridge as a function
%of fluid volume}
%\par\end{centering}
%\end{figure}
%
%
%\section{Capillary Bridge Model: Effect of hair diameter at constant fluid
%volume}
%
%Here, instead of scaling the fluid volume relative to the hair diameter,
%we now assume a fixed total fluid volume distributed equally among
%the $N$ hairs. Total fluid volume, $V_{total}=NV_{f}=2000$. Hair
%diameter is varied while keeping the total hair contact area constant.
%Length is in arbitrary units. Forces increase at a much smaller rate
%on decreasing diameter when compared to the case with self-similar
%scaling of fluid volume (Figure 8 in main text).
%
%\begin{figure}[H]
%\includegraphics{FigureS3-Effect_of_hair_size(constant_vol)}
%
%\caption{Normalised adhesion force of hairy pad system on a hydrophilic and
%hydrophobic substrate as a function of hair diameter ($D_{h}$), calculated
%from the capillary bridge model. The total adhesive fluid volume is
%fixed to 2000. Adhesion forces are calculated from minima of the respective
%force-distance curves. Negative force value represents attraction.
%The bubble's contribution to the net force for an \emph{underwater:
%bubble} contact is denoted by plus symbols. Bubble volume and pad
%diameter are kept fixed. All lengths are scaled relative to $D_{p}$.}
%\end{figure}

\section{Substrate characterization}

The surface chemistry of untreated glass (hydrophilic) and PFOTS-coated glass (hydrophobic) were characterized using dynamic contact angle measurements (Table \ref{tab:Contact-Angles}).
\begin{table}[H]
\centering{}%
\begin{tabular}{|c|c|c|c|}
\hline 
Substrate & Liquid & \ensuremath{\theta}\textsubscript{A} & \ensuremath{\theta}\textsubscript{R}\tabularnewline
\hline 
\hline 
\multirow{2}{*}{Glass} & Water & 63\ensuremath{\pm}5\textdegree{} & 20\ensuremath{\pm}2\textdegree{}\tabularnewline
\cline{2-4} \cline{3-4} \cline{4-4} 
 & n-Hexadecane & <10\textdegree{} & <10\textdegree{}\tabularnewline
\hline 
\multirow{2}{*}{PFOTS} & Water & 122\ensuremath{\pm}1\textdegree{} & 93\ensuremath{\pm}2\textdegree{}\tabularnewline
\cline{2-4} \cline{3-4} \cline{4-4} 
 & n-Hexadecane & 88\ensuremath{\pm}2\textdegree{} & 56\ensuremath{\pm}5\textdegree{}\tabularnewline
\hline 
\end{tabular}\caption{Dynamic contact angles (Mean \ensuremath{\pm} SD, n = 3) of Milli-Q water and n-hexadecane on the different test substrates. \label{tab:Contact-Angles}}
\end{table}

\section{Statistical comparison}

Two-way ANOVA test showed a significant effect of the \emph{Contact mode} (p=0.001, F=9.596, degrees of freedom=2) 
and \emph{Substrate} (p<0.001, F=36.231, degrees of freedom=1) categories on the single leg adhesion force measurements
of the ladybug beetle (\emph{Coccinella septempuctata}). Signifanct interaction between the above two categories was seen (p=0.001, F=10.551, degrees of freedom=2). Post-hoc analysis results are shown below (Table \ref{tab:Statistical-analysis}).
The uncorrected p-values and Common Language Effect Size (CLES) were
obtained from pair-wise Student t-test between A and B while
keeping the third parameter fixed (degrees of freedom=8 for each pair). p-values showing statistically
significant difference between A and B are in boldface. 
CLES represents the statistical proportion of samples under A with higher adhesion than under B. The condition
for statistical significance is based on the Bonferroni-corrected
critical p-value of 0.008.

\begin{table}[H]
\noindent \begin{centering}
\begin{tabular}{|>{\raggedright}m{0.15\linewidth}|>{\raggedright}m{0.15\linewidth}|>{\raggedright}m{0.15\linewidth}||>{\centering}m{0.1\linewidth}|>{\centering}m{0.1\linewidth}|>{\centering}m{0.1\linewidth}|}
\hline
Fixed variable & A & B & T & p-value & CLES\tabularnewline
\hline 
\hline 
In air & PFOTS & Glass & -0.053 & 0.959 & 0.48\tabularnewline
\hline
Underwater: bubble & PFOTS & Glass & 3.292 & 0.011 & 0.96\tabularnewline
\hline
Underwater: no bubble & PFOTS & Glass & 10.044 & 0.0 & 1.0\tabularnewline
\hline
PFOTS & In air & Underwater: bubble & 0.133 & 0.897 & 0.48\tabularnewline
\hline
PFOTS & In air & Underwater: no bubble & -0.224 & 0.828 & 0.48\tabularnewline
\hline
PFOTS & Underwater: bubble & Underwater: no bubble & -0.37 & 0.721 & 0.44\tabularnewline
\hline
Glass & In air & Underwater: bubble & 4.688 & 0.002 & 1.0\tabularnewline
\hline
Glass & In air & Underwater: no bubble & 11.341 & 0.0 & 1.0\tabularnewline
\hline
Glass & Underwater: bubble & Underwater: no bubble & 2.086 & 0.07 & 0.84\tabularnewline
\hline 
\end{tabular}
\par\end{centering}
\caption{Post-hoc t-test results for each combination of contact mode and substrate\label{tab:Statistical-analysis}}
\end{table}

\section{Capillary force due to an air bubble\label{subsec:Capillary-force-due}}

Capillary force of a single air bubble against a PFOTS-coated glass surface are compared
for two different volumes (Figure \ref{fig:Bubble_capillary_force}). The volumes correspond to the expected
range for the case of the trapped air bubble in a ladybug's pad. Here, the bubble
was pinned to a micropatterned PDMS substrate on the top. Approach-retract tests were performed
at 62.5 \textmu m s\protect\textsuperscript{-1} speed. The maximum
adhesion force of any of the bubble never exceeds 50 \textmu N,
significantly lower than the beetle's underwater adhesion to the same
substrate (> 400 \textmu N). Thus, the bubble's contribution
to adhesion in the ``\emph{underwater: bubble}'' contact of a ladybug's
pad should be negligible. Example measurement video is included in the supplementary data (Movie3).

\begin{figure}[H]
\begin{centering}
\includegraphics{FigureS4-Expt_bubble_force}\caption{Capillary force of the pinned bubble against a PFOTS-coated glass surface \label{fig:Bubble_capillary_force}}
\par\end{centering}
\end{figure}

\section{Supplementary video files}

\subsection*{Movie1}
Adhesion test recordings showing the three contact modes: \emph{in air},  \emph{underwater: bubble} and  \emph{underwater: no bubble} on a hydrophobic PFOTS-coated glass substrate. The two top panels of the video show the synchronous raw bottom-view and side-view recordings of the pad making contact with the substrate. The lower-left panel shows contact area extraction of the hairs with the surface via image processing and lower-right panel shows the corresponding temporal contact force and area data plot, with the data cursor synchronized with the other panels.

\subsection*{Movie2}
Adhesion test recording corresponding to the case of \emph{bad contact}, which occurred underwater on the PFOTS-coated glass substrate

\subsection*{Movie3}
Adhesion test recording of an air bubble (2nL volume) pinned to a microstructured PDMS on the top and making contact with a smooth PFOTS-coated glass substrate on the bottom.

\bibliography{references}

\end{document}